---
title: "R Notebook"
output: github_document
---

Mutating joins, add new variables to one data frame from matching observations in another.

Filtering joins, filter observations from one data frame based on whether or not they match an observation in the other table.

    semi_join(x, y) keeps all observations in x that have a match in y.
    anti_join(x, y) drops all observations in x that have a match in y.


Set operations treat observations as if they were set elements.

```{r}
library(tidyverse)
library(nycflights13)
library(Lahman)
library(babynames)
library(nasaweather)
library(fueleconomy)
library(ggplot2)
```

Exercises:

1.You would only need the flights table, with the origin and destination variables
2. The airports table and weather table can be merged by matching origin
3.It would align with the destinations of flights
4. Year, Month, Day, Hour, #_of_People. The keys would be time data.

Primary key identifies its own table, foreign key uniquely identifies and observation in another table

If a table contains multiple values that are the same....it isn't a KEY!!!
you can check

planes %>%
  count(tailnum) %>%
  filter( n> 1)
  
If a table doesn't seem to hold unique values anywhere, you can add your own key in. Use mutate() and row_number()

Exercises
```{r}
flights %>%
  mutate(flight_id = row_number()) %>%
  glimpse()

Batting %>%
  count(playerID, yearID, stint) %>%
  filter( n > 1)

babynames %>%
  count(year,sex, name)%>%
  filter(n > 1)

nasaweather::atmos %>%
 count(lat, long, year, month) %>%
  filter(n > 1)
  #Filter(n>1)

vehicles %>%
  count(id) %>%
  filter(n>1)
```
player id, year id, and stint

```{r}
flights2 <- flights %>%
  select(year:day, hour, origin, dest, tailnum, carrier)
#view(flights2)
flights2 %>%
  select(-origin, -dest)%>%
  left_join(airlines, by = "carrier")

x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3"
)

y <- tribble(
   ~key, ~val_y,
  1, "y1",
  2, "y2",
  4, "y3"
)

x %>%
  inner_join(y, by ="key")


```

inner join - matches observations whenever their keys are equal
outer join - Keeps observations that appear in both tables in at least 1 of the tables
  left join - keeps all in x of (x & y)
  right join - keeps all in y of (x & y)
  full join - keeps all observations in x and y
  



If there are duplicate keys, the result will be the cartiasian product of the results.

when joining by null the result is all variables that appear in both tables
if there are same named tables, but they mean different things specify by "value" and it will be added as a separate column

A named character vector e.g. "by = c("a" = "b"). This will match variable a in table x to variable b in table y. The variables from x will be used in the output. This means if you are trying to match destations of flights to a table with airpots and latitude and longitude you match by = c("dest" = "faa")

```{r}
flights %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE))%>%
  left_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(lon, lat, color = avg_delay)) +
  borders("state")+
  geom_point()+
  coord_quickmap()
```

```{r}
flights %>%
  left_join(airports, by = c("dest" = "faa"))
```

dplyr 	SQL
inner_join(x, y, by = "z") 	SELECT * FROM x INNER JOIN y USING (z)
left_join(x, y, by = "z") 	SELECT * FROM x LEFT OUTER JOIN y USING (z)
right_join(x, y, by = "z") 	SELECT * FROM x RIGHT OUTER JOIN y USING (z)
full_join(x, y, by = "z") 	SELECT * FROM x FULL OUTER JOIN y USING (z)

```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest
```



